Road traffic sound level estimation from realistic urban sound mixtures by Non-negative Matrix Factorization

### 2018-03-07 paper/gloaguenAppliedNMF.tex
proofread until db

\ml{la fin de ce par est a revoir car c'est le point crucial du papier. Bien expliquer le corpus du papier d'avant son usage et en quoi on passe ici a l'étape superieure.}

In this paper, the NMF framework is applied on a corpus of simulated sound scenes, generated based on annotated urban recordings, and whose realism is validated by a perceptual test. The NMF framework and its different optimizatio strategies are described in section \ref{part:nmf}. Next, the corpus of urban sound scenes is presented in section \ref{part:urban_scene}, from the sound database built-up to its validation through a perceptual test. The experimental protocol and the results are then presented and discussed in section \ref{part:expProtocol} and \ref{part:results}.

@@

The choice of the dimensions is often made such as $F\times K + K \times N < F \times N$ so that NMF can be a low rank approximation. This condition however is not mandatory. When an audio file is considering, $\mathbf{V}$ is usually considered as the magnitude spectrogram obtained by a Short-Time Frequency Transform, $\mathbf{W}$ includes audio spectra and $\mathbf{H}$ is equivalent to the temporal evolution of each spectrum, see Figure \ref{fig:exampleNMF}. Because of the non-negativity constraint, only additive combinations between the elements of $\mathbf{W}$ are considered. The dictionary is then composed of elementary elements providing a part based representation.

\begin{figure}[t]
\includegraphics[width=.9\linewidth]{./figures/schema_introduction_nmf.pdf}
\caption{NMF for an audio sample with 3 elements ($K$ = 3): passing car (a), car horn (b) and whistling bird (c).}
\label{fig:exampleNMF}
\end{figure}

@@

$d_{\beta}(x\vert y)$ is mainly chosen as a $\beta$-divergence \cite{fevotte_algorithms_2011}, a sub-classes belonging to the Bregman divergences \cite{hennequin_beta-divergence_2011} which include 3 specific divergence and distance calculations: the Euclidean distance (eq. \ref{eq:def_distEUC}), the Kullback-Leibler divergence (eq. \ref{eq:def_divKL}) and the Itakura-Sa\"{i}to divergence (eq. \ref{eq:def_divIS}):

@@

The minimization problem (\ref{eq:min-D-WH}) is solved iteratively by updating the form of matrices $\mathbf{W}$ and $\mathbf{H}$. Different algorithms such as Alternating Least Square Method \cite{cichocki_regularized_2007} or Projected Gradient \cite{lin_projected_2007} have been proposed. The most commonly algorithm used, and the chosen method here, is Multiplicative Update \cite{lee_algorithms_2000} as it ensures non-negative results and the convergence of the results \cite{fevotte_algorithms_2011}.

\subsection{Supervised NMF}
The most easiest case of NMF is the one where the sound sources can be known \textit{a priori} and $\mathbf{W}$ can be built directly from audio samples. It leads to \textit{supervised} NMF (SUP-NMF). $\mathbf{H}$ is then the only matrix to determine and is updated at every iteration (eq. \ref{eq:updateH}) \cite{fevotte_algorithms_2011}.

@@

Here, in an urban context, the sound sources are known and their audio samples can be obtained to learn $\mathbf{W}$, see section \ref{part:dictionary_building}. As the position of each element is indexed, the traffic source separation from the other sound sources is made by extracting, from the dictionary and the activation matrix, the related elements:

@@

\subsection{Semi-supervised NMF}
The main issue with the supervised approach is the limit imposed by $\mathbf{W}$: to be successful, all the acoustical sources must be considered in it which is not possible in an urban environment. As the main source that interests us can be known \textit{a priori}, semi-supervised NMF (SEM-NMF) \cite{lee_semi-supervised_2010} is considered to better take into account the interfering sound sources. This approach proposes to decompose $\mathbf{W}_{F \times (K+J)}$ with two distinctive matrices: $\mathbf{W} = \left[ \mathbf{W_s}~\mathbf{W_r} \right]$ where $\mathbf{W_s}_{F \times K}$ is a fixed part of $\mathbf{W}$ composed of audio spectra and $ \mathbf{W_r}_{F \times J}$, a mobile part which is updated, see eq. \ref{eq:W_r_SS}. Thus it is possible to include elements not present in $\mathbf{W_s}$. The dimension of $\mathbf{W_r}$ is set up as $J << K$ in order to best considered the sound source present in $\mathbf{W_s}$. $\mathbf{H}$ is then also decomposed in two matrices, $\mathbf{H}_{(K+J) \times N} = \genfrac[]{0pt}{0}{\mathbf{H_s}}{\mathbf{H_r}}$. The eq. \ref{eq:nmf} becomes

@@

This approach as the advantage, with the add of the mobile part $\mathbf{W_r}$, to bring more flexibility and then to be more adaptive to the different urban sound environments. Applications of SEM-NMF can be found for musical \cite{weninger2012supervised, kitamura_music_2014} and speech content \cite{joder2012real, mysore2011non}.

@@

We propose a third approach based on the unsupervised NMF framework: Threshold Initialized NMF (TI-NMF). Usually, in unsupervised NMF, the dictionary is initiated randomly when there is no \textit{prior} knowledge on the sound sources present. Here, as the target sound source is known and the spectra are available, an initial dictionary, $\mathbf{W_0}$, is designed and then updated alternatively with $\mathbf{H}$,

@@

where $\mathbf{w}$ is a $k$ element of $\mathbf{W}$ of $F \times 1$ dimension. When $D_{\theta}\left(\mathbf{w_0} \Vert \mathbf{w'} \right)$=1, the element $k$ from $\mathbf{W'}$ is exactly similar than in $\mathbf{W_0}$. If $D_{\theta}\left(\mathbf{w_0} \Vert \mathbf{w'} \right)$=0, the element is fully different. Next, the similarities are sorted in descending order. The extraction of traffic elements in $\mathbf{W'}$ is carried out by a hard thresholding method \cite{donoho1994threshold}. It consists in weighting in a binary way the traffic elements $\mathbf{W'}$ according to $D_{\theta}\left(\mathbf{w_0} \Vert \mathbf{w'} \right)$ and a threshold value such as:

@@

These methods are applied on simulated sound scenes in order to compare the estimated sound levels with the exact solutions. In \cite{gloaguen2018Estimation}, the sound corpus was composed of a mix of traffic samples with specific sound classes (\textit{alert, animals, climate, human, mechanics, transportation}). Here, in order to implement this method in embedded sensors, a new more realistic sound corpus is generated based on urban recordings mixing all kind of sound sources.

>>>>

\ml{la fin de ce par est a revoir car c'est le point crucial du papier. Bien expliquer le corpus du papier d'avant, son usage, les questions qui restent en suspens et en quoi on passe ici a l'étape superieure en y repondant.}

In this paper, the NMF framework is applied on a corpus of simulated sound scenes, generated based on annotated urban recordings, and whose degree of realism is validated through a perceptual test. The NMF framework and its different optimization strategies are described in section \ref{part:nmf}. Next, the corpus of urban sound scenes is presented in section \ref{part:urban_scene}, from the sound database built-up to its validation. The experimental protocol and the results are then presented and discussed in section \ref{part:expProtocol} and \ref{part:results}.

@@

The choice of the dimensions is often made such as $F\times K + K \times N < F \times N$ so that NMF can be a low rank approximation. This condition however is not mandatory. When  applying NMF to audio data, $\mathbf{V}$ is usually considered as the magnitude spectrogram obtained by a Short-Time Frequency Transform, $\mathbf{W}$ includes audio spectra and $\mathbf{H}$ is equivalent to the temporal activation of each spectrum, see Figure \ref{fig:exampleNMF}. Because of the non-negativity constraint, only additive combinations between the elements of $\mathbf{W}$ are considered. %The dictionary is then composed of elementary elements providing a part based representation.

\begin{figure}[t]
\includegraphics[width=.9\linewidth]{./figures/schema_introduction_nmf.pdf}
\caption{NMF decomposition of an audio spectrogram composed of 3 elements ($K$ = 3): passing car (a), car horn (b) and whistling bird (c).}
\label{fig:exampleNMF}
\end{figure}

@@

$d_{\beta}(x\vert y)$ is usually chosen as a $\beta$-divergence \cite{fevotte_algorithms_2011}, a sub-classes belonging to the Bregman divergences \cite{hennequin_beta-divergence_2011} which include 3 specific divergence calculations: the Euclidean distance (eq. \ref{eq:def_distEUC}), the Kullback-Leibler divergence (eq. \ref{eq:def_divKL}) and the Itakura-Sa\"{i}to divergence (eq. \ref{eq:def_divIS}):

@@

The minimization problem (\ref{eq:min-D-WH}) is solved iteratively by updating the form of matrices $\mathbf{W}$ and $\mathbf{H}$. Different algorithms such as Alternating Least Square Method \cite{cichocki_regularized_2007} or Projected Gradient \cite{lin_projected_2007} have been considered. The most commonly used algorithm is the Multiplicative Update \cite{lee_algorithms_2000}. The latter method is chosen here, as it ensures non-negative results and the convergence of the results \cite{fevotte_algorithms_2011}.

\subsection{Supervised NMF}
The most easiest case of NMF is the one where the sound sources can be known \textit{a priori} and $\mathbf{W}$ can be built directly from audio samples. It leads to \textit{supervised} NMF (SUP-NMF). $\mathbf{H}$ is then the only matrix to estimate and is updated at every iteration (eq. \ref{eq:updateH}) \cite{fevotte_algorithms_2011}.

@@

Here, in an urban context, if the sound sources are known, their audio samples can be obtained to learn $\mathbf{W}$, see section \ref{part:dictionary_building}. As the position of each element is indexed, the traffic source separation from the other sound sources is made by extracting, from the dictionary and the activation matrix, the related elements:

@@

\subsection{Semi-supervised NMF}

The main issue with the supervised approach is the representational limit imposed by a fixed $\mathbf{W}$. To be completely successful, all the acoustical sources must be considered in it which is not possible in an urban environment. However, in our application setting, the main source of interest, \textit{i.e.} the traffic, can be known \textit{a priori} \ml{manque une phrase} semi-supervised NMF (SEM-NMF) \cite{lee_semi-supervised_2010} is considered to better take into account the interfering sound sources. To do so, $\mathbf{W}_{F \times (K+J)}$ is decomposed into two distinctive matrices: $\mathbf{W} = \left[ \mathbf{W_s}~\mathbf{W_r} \right]$ where $\mathbf{W_s}_{F \times K}$ is a fixed part of $\mathbf{W}$ composed of audio spectra and $ \mathbf{W_r}_{F \times J}$, a mobile part which is updated, see eq. \ref{eq:W_r_SS}. Thus it is possible to include elements not present in $\mathbf{W_s}$. The dimension of $\mathbf{W_r}$ is set up as $J << K$ in order to best considered the sound source present in $\mathbf{W_s}$. $\mathbf{H}$ is then also decomposed in two matrices, $\mathbf{H}_{(K+J) \times N} = \genfrac[]{0pt}{0}{\mathbf{H_s}}{\mathbf{H_r}}$. The eq. \ref{eq:nmf} becomes

@@

The addition of the mobile part $\mathbf{W_r}$ gives more flexibility. The representational capability is increased, thus the approach is more adaptive to the different urban sound environments. Applications of SEM-NMF can be found for musical \cite{weninger2012supervised, kitamura_music_2014} and speech content \cite{joder2012real, mysore2011non}.

@@

To allow even more flexibility while still considering prior knowledge of the source of interest, we propose a third approach based on the unsupervised NMF framework: Threshold Initialized NMF (TI-NMF). Usually, in unsupervised NMF, the dictionary is initiated randomly when there is no \textit{prior} knowledge on the sound sources present. Here, as the target sound source is known and the spectra are available, an initial dictionary, $\mathbf{W_0}$, is designed and then updated alternatively with $\mathbf{H}$,

@@

where $\mathbf{w}$ is a $k$ element of $\mathbf{W}$ of $F \times 1$ dimension. When $D_{\theta}\left(\mathbf{w_0} \Vert \mathbf{w'} \right)$=1, the element $k$ from $\mathbf{W'}$ is equal to the in $\mathbf{W_0}$ \ml{pas clair, a quoi sert le k ?}. If $D_{\theta}\left(\mathbf{w_0} \Vert \mathbf{w'} \right)$=0, the element is fully different. Next, the similarities are sorted in descending order \ml{le sort n'est pas necessaire, le seuil suffit}. The extraction of traffic elements in $\mathbf{W'}$ is carried out by a hard thresholding method \cite{donoho1994threshold}. It consists in weighting in a binary way the traffic elements $\mathbf{W'}$ according to $D_{\theta}\left(\mathbf{w_0} \Vert \mathbf{w'} \right)$ and a threshold value such as:

\ml{on pourrait tout a fait imaginer l'ajout d'élément de dictionaire random pour se rapprocher de la semi.}

@@

\ml{faire un par resumant les prop des trois methodes avec un accent sur l'utilisation de la connaissance a priori.}

These methods are applied on simulated sound scenes in order to compare the estimated sound levels with the exact solutions. In \cite{gloaguen2018Estimation}, the sound corpus was composed of a mix of traffic samples with specific sound classes (\textit{alert, animals, climate, human, mechanics, transportation}). Here, in order to implement this method in embedded sensors, a new and more realistic sound corpus is generated based on urban recordings mixing many kind of sound sources.
